---
title: "Data 607 - Project 2"
author: "SG"
output: github_document
always_allow_html: yes
---

```{r}
library(magrittr)
library(stringr)
library(dplyr)
library(tidyr)
library(zoo)
library(ggplot2)
library(knitr)
library(DT)
```

#Project 2 - Perform analysis on wide data sets  
We will analyze three different wide data sets:  
1. Salaries by College Type - https://www.kaggle.com/wsj/college-salaries/version/1#salaries-by-college-type.csv This was found from the discussion board, post title: Salaries by College Degree/Type/Region  
2. Graduate Admissions - https://www.kaggle.com/mohansacharya/graduate-admissions This was a data set I found on my own.    
3. Employee Reviews - https://www.kaggle.com/petersunga/google-amazon-facebook-employee-reviews This was found the from the discussion board, post title: Employee Reviews at top tech companies      
These data sets have a large amount of data, so we will take sample of n size = 200, from each data set.  

## Salaries by College Type  
We will analyze salaries by college type. It will be interesting to see how the college type is correlated with salaries, at:
1. Mid Career Median Salaries by College Type and by Percentiles
2. Starting Salaries by College Type   

### Sample the Raw Data
We get the sample of raw data for Salaries by College Type.  
```{r}
collegeTypeData <- read.csv("Data/salaries-by-college-type.csv")
collegeTypeDf <- sample_n(collegeTypeData, 200)
#collegeTypeDf <- na.omit(collegeTypeDf)
datatable(collegeTypeDf, options = list(pageLength = 5))
```

### Reduce width of data  
We reduce the width of the data. The following columns make the data set wide:  
* Mid-Career 10th Percentile Salary  
* Mid-Career 25th Percentile Salary  
* Mid-Career 75th Percentile Salary  
* Mid-Career 90th Percentile Salary  
We transform these columns into two different columns: MidCareerMedianPercentile and PercentileSalary, to view salaries by their Mid Career Median Percentiles. We also reduce the string in the MidCareerMedianPercentile column to a number. The data is now cleaner to analyze:  
```{r}
# Reduce width of Percentile columns
collegeTypeDf <- collegeTypeDf %>% gather(key = MidCareerMedianPercentile, value = PercentileSalary, Mid.Career.10th.Percentile.Salary:Mid.Career.90th.Percentile.Salary)
collegeTypeDf[collegeTypeDf == "N/A"]  <- 0

# Convert all Salary columns from string to numeric types
collegeTypeDf$Starting.Median.Salary <- as.numeric(gsub('[$,]', '', collegeTypeDf$Starting.Median.Salary))
collegeTypeDf$Mid.Career.Median.Salary <- as.numeric(gsub('[$,]', '', collegeTypeDf$Mid.Career.Median.Salary))
collegeTypeDf$PercentileSalary <- as.numeric(gsub('[$,]', '', collegeTypeDf$PercentileSalary))

# Apply regex to get Percentile numbers
collegeTypeDf$MidCareerMedianPercentile <- sub("Mid.Career.", "\\1", collegeTypeDf$MidCareerMedianPercentile)
collegeTypeDf$MidCareerMedianPercentile <- sub(".Percentile.Salary", "\\1", collegeTypeDf$MidCareerMedianPercentile)

datatable(collegeTypeDf, options = list(pageLength = 5))
```
  
###Analysis

**1. What was the average salary for each Median Career Percentile?**  
```{r}
avgByPercentile <- aggregate(x=collegeTypeDf$PercentileSalary,
          by=list(collegeTypeDf$School.Type,collegeTypeDf$MidCareerMedianPercentile),
          FUN=mean)
colnames(avgByPercentile) <- c("CollegeType", "MidCareerMedianPercentile", "AverageSalary")
datatable(avgByPercentile, options = list(pageLength = 5))
```
+ **Grouped Bar Plot of the above averages**  
We see that Ivy Leage colleges had the highest average salaries for Mid Career Median Percentiles. Second highest salaries were for State colleges.  

```{r}
ggplot(avgByPercentile, aes(CollegeType, AverageSalary)) + geom_bar(aes(fill = MidCareerMedianPercentile),
   width = 0.4, position = position_dodge(width=0.5), stat="identity") +  
   theme(legend.position="top", legend.title = 
   element_blank(),axis.title.x=element_blank(), 
   axis.title.y=element_blank())
```

**Analyzing Starting Salaries**  
**College Types with Starting Salaries > $60,000**  

Here, we see only Engineering and Ivy League schools with starting salaries above $60,000.    

```{r}
#collegeTypeDf$Starting.Median.Salary <- as.numeric(collegeTypeDf$Starting.Median.Salary)
greaterThanSixtyK<- collegeTypeDf %>% select(School.Name, School.Type, Starting.Median.Salary) %>% filter(Starting.Median.Salary > 60000)
datatable(greaterThanSixtyK, options = list(pageLength = 5))
```
  
Here we see a bar plot of the above table. Engineering schools are displayed to have higher starting salaries than Ivy League schools.  
```{r}
ggplot(data = greaterThanSixtyK, aes(x = School.Type, y = Starting.Median.Salary)) + 
       geom_bar(stat = 'identity', position = 'dodge', fill="LightBlue")
```  
  

## Graduate Admissions   
We will analyze graduate admissions using by:
* Letters of Recommendation and State of Purpose by Chances of Admission  

### Sample the Raw Data
We get the sample of raw data for Graduate Admissions.    
```{r}
admissionsData <- read.csv("Data/Admission_Predict.csv")
admissionsDf <- sample_n(admissionsData, 200)
datatable(admissionsDf, options = list(pageLength = 5))
```

### Reduce width of data  
We reduce the width of the data by:
1. Combining columns of Letter of Recommendation and State of Purpose, and a separate column for their values. We can combine these two columns because they're both based on a rating system of out of 5.  
```{r}
admissionsDf <- admissionsDf %>% gather(key = LetterType, value = LetterStrength, SOP:LOR)
admissionsDf$Chance.of.Admit <- as.numeric(admissionsDf$Chance.of.Admit)
admissionsDf$LetterStrength <- as.numeric(admissionsDf$LetterStrength)
datatable(admissionsDf, options = list(pageLength = 5))
```

### Analysis  

#### Statement of Purpose and Letters of Recommendation by Chances of Admission  

Here we view a bar plot of the Chances of Admission by State of Purpose and Letters of Recommendation Strengths. We see that the higher the Strengths are, like by strengths of 4 or 5, the higher the chance of admission is. Chances of admission seemed to fair higher with Letters of recommendation than with Statements of Purpose, when strengths were lower, at about 1-3.    
```{r}
ggplot(admissionsDf, aes(LetterStrength, Chance.of.Admit)) + geom_bar(aes(fill = LetterType),
   width = 0.4, position = position_dodge(width=0.5), stat="identity") +  
   theme(legend.position="top", legend.title = 
   element_blank(),axis.title.x=element_blank(), 
   axis.title.y=element_blank())
```

Below we see the that the average strengths of Letters of Recommendations and Statements of Purpose seemed to be almost the same.  
```{r}
aggregAdmissions <- aggregate(x=admissionsDf$LetterStrength,
          by=list(admissionsDf$LetterType),
          FUN=mean)
colnames(aggregAdmissions) <- c("Letter Type", "Letter Strength")
datatable(aggregAdmissions)
```

####Chance of Admission by GPA  
Below we observe the Chance of Admission vs. GPA. We see an overall positive linear correlation between the Chance of Admission and GPA, which makes sense.  

```{r}
ggplot(admissionsDf, aes(CGPA)) + 
  geom_line(aes(y = Chance.of.Admit, colour = "Admission Chance"))
```

## Employee Reviews  
It is interesting to see correlations between companies and employee reviews. Here, we will be analyzing data on employee reviews for Google, Amazon, Facebook, Apple, and Microsoft, by the following:  
* Summary of Ratings by Company  

### Sample the Raw Data
We get the sample of raw data for Employee Reviews.  

```{r}
emplReviewData <- read.csv("Data/employee_reviews.csv")
emplReviewDf <- sample_n(emplReviewData, 200)
datatable(emplReviewDf, options = list(pageLength = 5))
```

###Reduce width of the data  

There are several columns for ratings, that are measured 1-5:  
* Work/Life Balance Rating  
* Culture and Values Rating  
* Career Opportunities Rating  
* Comp & Benefits Rating  
* Senior Management Rating  

These columns are repetitive. So we clean this data by combining these columns into a column called RatingType and a column for each value called Rating.  

```{r}
emplReviewDf <- emplReviewDf %>% gather(key = RatingType, value = Rating, 11:15)
emplReviewDf$overall.ratings <- as.numeric(emplReviewDf$overall.ratings)
emplReviewDf$Rating <- gsub("[[:space:]]*","0",emplReviewDf$Rating)
emplReviewDf$Rating <- as.numeric(emplReviewDf$Rating)
datatable(emplReviewDf)
```

###Analysis   

####Average Rating Types by Company  
We analyze overall ratings by employees between companies. It would be interesting to see which companies received the highest type of ratings. Please note, due to not cleaning up a few blanks or na's in the Ratings data, we may not see some values.  

```{r}
  avgRatingByCompany <- aggregate(x=emplReviewDf$Rating,
          by=list(emplReviewDf$company,emplReviewDf$RatingType),
          FUN=mean)
colnames(avgRatingByCompany) <- c("Company", "RatingType", "Rating")
datatable(avgRatingByCompany, options = list(pageLength = 5))
```

Here, we see a summary bar plot of ratings and their types by company. We see that Facebook and Netflix had relatively higher ratings. Please note, due to not cleaning up a few blanks or na's in the Ratings data, we may not see some values.:    

```{r}
ggplot(avgRatingByCompany, aes(Company, Rating)) + geom_bar(aes(fill = RatingType),
   width = 0.4, position = position_dodge(width=0.5), stat="identity") +  
   theme(legend.position="top", legend.title = 
   element_blank(),axis.title.x=element_blank(), 
   axis.title.y=element_blank())
```
