---
title: "DATA 607 - Assignment 3"
author: "SG"
output: github_document
---

```{r}
  library(stringr)
```
## 3. Copy the introductory example. The vector name stores the extracted names.  
```{r}
  name <- c("Moe Szyslak", "Burns, C. Montgomery", "Rev. Timothy Lovejoy", "Ned Flanders", "Simpson, Homer", "Dr. Julius Hibbert")
  name
```
**Problem 1.** Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.  
First we split each row in the vector by the comma followed by the space:  
```{r}
  splitNames <- str_split(name, ", ", simplify = TRUE)
```
Next, we concatenate the first name with the last name, separated by a space:  
```{r}
  str_c(splitNames[,2], " ", splitNames[,1])
```

**Problem 2.** Construct a logical vector indicating whether a character has a title(i.e., Rev. and Dr.)  
```{r}
  name
  isTitledName <- str_detect(string = name, pattern = "^\\w+[.][[:space:]](\\w+[[:space:]]\\w+)*")
  isTitledName
```
**Problem 3.** Construct a logical vector indicating whether a character has a second name.  
```{r}
  name
  hasSecondName <- str_detect(string = name, pattern = "[[:alpha:]]+([,]|[.])?[[:space:]][[:alpha:]]+([.]|[,]^)?[[:space:]]([[:alpha:]]+([.]|[,]^)?([[:space:]])?)*")
  hasSecondName
```
## 4. Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.  
**Problem 1.**"[0-9]+\\$"
One or more digits that end in $
```{r}
  example <- c("12345\\", "0$", "123456$", "$")
  regexExample <- unlist(str_extract_all(example, pattern="[0-9]+\\$", simplify=TRUE))
  regexExample
```
**Problem 2.** \\b[a-z]{1-4}\\b  
Lowercase letters, of a minumum length of 1 to a maximum length of 4.  
```{r}
example <- c("abcd", "helloWorld", "abc", "1,4", "e", "abcdefg")
regexExample <- str_extract_all(example, pattern = "\\b[a-z]{1,4}\\b", simplify=TRUE)
regexExample
```
**Problem 3.** .*?\\.txt$  
Any string of any characters that ends in .txt  
```{r}
  example <- c(".txt", "test.txt", "abc", "12345&*@&#$.txt", "test.TXT", " .txt", ".txttxt", ".txtHello")
  regexExample <- str_extract_all(example, ".*?\\.txt$", simplify=TRUE)
  regexExample
```

**Problem 4.** \\d{2}/\\d{2}/\\d{4}  
This regex takes in a date format. String: two digits followed by a slash, followed by two digits, followed by another slash, followed by 4 digits
```{r}
  example <- c("02/17/2019", "12/", "00/00/0000", "abc/a/aa", "1/1/1111")
  regexExample <- str_extract_all(example, "\\d{2}/\\d{2}/\\d{4}", simplify=TRUE)
  regexExample
```

**Problem 5.** <(.+?)>.+?</\\1>
This regex seems to be for HTML formats. The string it takes is: any set of characters between < and >, followed by any set of characters, followed by <, the same set of characters that were previously between < and >, a slash, and then >  
```{r}
  example <- c("<b>bold text</b>", "<a>?</a>", "<b>Work</a>", "<#$%&*>Hello</#$%&*>", "<h1></h1>")
regexExample <- str_extract_all(example, "<(.+?)>.+?</\\1>", simplify=TRUE)
regexExample
```

##9. Extra Credit: clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr  
This message reads as CONGRATULATIONS YOU ARE A SUPER NERD!

```{r}
  extraCredit <- (unlist(str_extract_all("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr", pattern="[A-Z]", simplify=TRUE)))
  extraCredit
```

