---
title: "Assignment5"
author: "SG"
date: "3/3/2019"
output: html_document
---
## Load libraries:  
- tidyr tidies the data  
- dplyr transforms the data

```{r}
 library(tidyr)
  library(dplyr)
library(ggplot2)
```

## Read the .CSV file:  

```{r}
fileData <- read.csv(file="//Users/suma/Desktop/CUNY SPS - Masters Data Science/Data 607/Assignment5/ArrivalDelays.csv", header = T)
head(fileData)
```

## Tidy the data using tidyr:  
Here, we use the gather method. This is where we have one row per case, a column for each variable, and a cell for each value. We put the data's columns into rows. In this way, we are making the data from wide to long. 

```{r}
fileData <- gather(fileData, Destination, ArrivalDelay, 3:7)
names(fileData) <- c("Airline", "Status", "Destination", "ArrivalDelay")

fileData
```

We want to further tidy the data by making it wide again, using the spread function, to split the Status column's rows into two columns: On Time and Delayed, with values being the numbers from the ArrivalDelay column. Note: I've been getting an error in the code for the spread function so I didn't spread this data in time.

```{r}
  #fileData <- spread(fileData, Status, ArrivalDelay, 3:7)
#fileData
```

## Transform the data using dplyr:  
Here, we filter the data as an example of data transformation. We filter by the destination of Phoenix.  
```{r}
  filter(fileData, Destination == 'Phoenix')
```

## Statistical Analysis:  
Here we analyze our data. There are some summary useful statistics for Arrival Delays. For the plot: It seems ggplot works best for long data, not wide data, which is why the graph below is not a meaningful data analysis. As mentioned earlier, with the coding error, I did not spread this data.  

```{r}
  summary(fileData)
  ggplot(fileData, aes(Airline)) + 
  geom_line(aes(y = Airline, colour = "Airline")) + 
  geom_line(aes(y = ArrivalDelay, colour = "ArrivalDelay"))
```


